---
title: "Programming in Base R"
format: html
editor: visual
---

# Task 1: Basic Vector practice

## 1. Create two vectors for the pre-treatment and post treatement measurements.

```{r}
PreTreatment <-c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

AfterTreatment <-c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

PreTreatment
AfterTreatment

paste("Subject", 1:20, sep = "_")

mydf <- data.frame(Subjects, PreTreatment, AfterTreatment)

```

## 2. Assign names to the vector elements.
```{r}
paste("Subject", 1:20, sep = "_")

```

## 3. Calculate change in blood pressure between treatements.
```{r}
TreatmentDiffs <- (AfterTreatment-PreTreatment) 

TreatmentDiffs
```
## 4. Calculate the average decrease in blood pressure across all patients. 
```{r}
mean(TreatmentDiffs)
```
## 5. Determine which patients experienced a decrease in blood pressure.
```{r}

```


# Task 2
## Create a data frame object with four columns corresponding to the above data. 
```{r}
mydf <- data.frame(Subjects, PreTreatment, AfterTreatment, TreatmentDiffs)
```

