---
title: "Programming in Base R"
format: html
editor: visual
---

# Task 1: Basic Vector practice

## 1. Create two vectors for the pre-treatment and post treatement measurements.

```{r}
PreTreatment <-c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

AfterTreatment <-c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

PreTreatment
AfterTreatment

```

## 2. Assign names to the vector elements.

```{r}
Subject <- paste("Subject", 1:20, sep = "_")
Subject

```

## 3. Calculate change in blood pressure between treatements.

```{r}
TreatmentDiffs <- (AfterTreatment-PreTreatment) 

TreatmentDiffs
```

## 4. Calculate the average decrease in blood pressure across all patients.

```{r}
mean(TreatmentDiffs)
```

## 5. Determine which patients experienced a decrease in blood pressure.

```{r}
Subject[which(AfterTreatment < PreTreatment)]
```

## 6. Subset the vector of differences to only return those that have a positive change

```{r}
TreatmentDiffs [which(AfterTreatment < PreTreatment)]
```

## 7. Calculate the average decrease in blood pressure.

```{r}
mean(TreatmentDiffs[which(AfterTreatment < PreTreatment)])
```

# Task 2

## 1. Create a data frame object with four columns corresponding to the above data.

```{r}
TreatmentsDF <- data.frame(Subject, PreTreatment, AfterTreatment, TreatmentDiffs)
TreatmentsDF
```

## 2. Return only rows where TreatmentDiffs column is negative.

```{r}
TreatmentsDF[which(TreatmentsDF$TreatmentDiffs < 0), ]
```

## 3. Add a new column to the data frame corresponding to TRUE if the AfterTreatment BP is less than 120.

```{r}
TreatmentsDF |> 
  mutate(BPUnder120 = TreatmentsDF$AfterTreatment < 120)
```

## 4. Print the data frame out nicely in the final documnet.

```{r}

```

# Task 3

## 1. Create a new data frame with the placebo experiment.

```{r}
PlaceboBefore <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
PlaceboAfter <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
PlaceboDiffs
```
